How to Avoid OOTA Without Really Trying

### 2025-03-07 paper/rebuttal-mods.txt

x	Be less condescending towards Mark Batty.  Maybe
	s/is persisting with/is working on/.

o	Add a Section 1.3 ("Our Approach"?) that summarizes the
	rfe/sdep/non-sdep-dep approach and emphasizes proving
	OOTA cannot happen as opposed to an exact delineation
	between sdep and non-sdep.

o	Add citations to later sections that discuss litmus tests
	that are similar to those in the literature.

o	Consider adding a 2.2.x section that discusses the difficulties
	inherent in computing sdep for function calls, especially when
	the function will be part of a dynamic library that has not yet
	been written.  (This addresses Reviewer #72F's call for handling
	examples in Section 5 of [5].)

@@

x	In section 6.1, define "witness" informally as "another execution
	in which the load(s) obtain a different value but all else is
	the same as far as possible, and yet the store acts differently".

o	Add diagrams illustrating what the various components of sections
	6.1 and 6.3 refer to. (Reviewer #72B)

x	On page 9, "While it possible".  [ Since fixed by Alan. ]

x	In Section 7.4 ("Merging Quasi-Volatile Loads"), change focus
	from intuitive notions to a community-process approach for
	determining what optimizations are legitimate.  Math cannot
	handle this.  We avoid the problem by looking at the compiler
	implementation, which is downstream of the community process.

### 2025-03-07 paper/ootavacuous.tex
ootavacuous: Add citations requested by Reviewer #72F
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

All OOTA researchers owe a debt of gratitude to the foundational work led
by William Pugh, represented by the infamous Java-language
``Causality Test Cases''.\footnote{
	\url{http://www.cs.umd.edu/~pugh/java/memoryModel/unifiedProposal/testcases.html}.}
% %%% Eventually cite P3064.

@@

Moiseenko et al.~avoid cycles in po $\cup$
rf~\cite{EvgeniiMoiseenko2025reexecution}, which might permit
more optimizations than forcing prior relaxed loads to be ordered
before subsequent relaxed stores, but which the authors characterize
as ``(almost) a per-execution model''.
It remains to be seen whether this is close enough to a per-execution
model to be accepted by C++ implmenters.
\emph{I am not fully confident of my characterization of this paper.}

@@

Therefore, this paper's contribution is to show how these constraints
prevent real-world C++ implementations from exhibiting OOTA behavior,
and not with minimal cost~\cite{PiezhaoOu2018OOTApapl}, but rather with
no cost at all.

>>>>

All OOTA researchers owe a debt of gratitude to the foundational work led
by William Pugh, represented by the infamous Java-language
``Causality Test Cases''\footnote{
	\url{http://www.cs.umd.edu/~pugh/java/memoryModel/unifiedProposal/testcases.html}.}~\cite{JeremyManson2005JavaMemoryModel}.
The OOTA problem has been open for more than 20~years and has been shown
to be quite difficult~\cite{MarkBatty2015OOTAisHard}.
% %%% @@@ Eventually cite P3064.

@@

Moiseenko et al.~avoid cycles in po $\cup$
rf~\cite{EvgeniiMoiseenko2025reexecution}, which might permit
more optimizations than forcing prior relaxed loads to be ordered
before subsequent relaxed stores, but which the authors characterize
as ``(almost) a per-execution model''.
It remains to be seen whether this is close enough to a per-execution
model to be accepted by C++ implmenters.
\emph{@@@ I am not fully confident of my characterization of this paper.}

@@

Therefore, this paper's contribution is to show how these
constraints prevent real-world C++ implementations doing per-thread
analysis from exhibiting OOTA behavior, and not with minimal
cost~\cite{PiezhaoOu2018OOTApapl}, but rather with no cost at all.
\emph{@@@ Are these the right qualifiers?}
